---
title: "ANOVA_Main_version"
author: "Galen O'Shea-Stone"
date: "8/26/2020"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(MetaboAnalystR)
library(tidyverse)
#library(reshape)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(dplyr)
#library(hrbrthemes)
#library(viridis)
#library(data.table)
library(plyr)
library(caret)
library(gplots)
library(ggsignif)
library(cowplot)
#library(ggpubr)

```

```{r, echo = FALSE, include=FALSE}
## This chunk reads in raw .csv performs scaling, normalization and log transform then outputs all the data back into the working directory
mSet <- InitDataObjects("conc", "stat", paired = FALSE)
mSet <- Read.TextData(mSet, "EXAMPLE_DATA.csv", "rowu", "disc")
#change from rowu to colu for column data vs row 
#rowp vs colp for paired analysis 

#Perform data processing
mSet <- SanityCheckData(mSet)
mSet <- ReplaceMin(mSet);
mSet <- PreparePrenormData(mSet)
mSet <- Normalization(mSet, "none", "LogNorm", "AutoNorm", ref= NULL, ratio = FALSE, ratioNum = 20)
mSet <- PlotNormSummary(mSet, "norm_O_", "png",72, width = NA)
mSet <- PlotSampleNormSummary(mSet, "snorm_O_", "png",72, width = NA)
mSet <-SaveTransformedData(mSet)


#after this is done, there will be a new files in the working directory 
#important one moving forward is titled data_normalized.csv
```


```{r}
mSet.1 <- ANOVA.Anal(mSet, nonpar = FALSE, thresh = 0.05, post.hoc = "tukey", all_results = FALSE)
#preforms ANOVA, change parameters above       
#see note below about non-parametric tests

#mSet.1

ANOVA.data <- read.csv("anova_posthoc.csv", sep=",")
#note: if doing non-parametric then need to change line above to "kw_posthoc.csv"



norm.data <- mSet$dataSet$norm
#extract the normalized data from the large list created in the preprocessing

status <- mSet$dataSet$cls


x <-  mSet$dataSet$norm
norm.data <- add_column(x, status, .before = T)
#combines the status set into the data for differentiation 

signif.values <- read.csv("anova_posthoc.csv")
#gathers list of significant metabolites from volcano analysis
s.v <- as.factor(signif.values$X)
#changes this for later use
```
```{r}

#this chunk will create the ANOVA plot(s)

###NOTE: occasionally this will cause an error with "s.v" portion. If this happens just clear environment AND restart R 
#probably  an issue with interaction with another package/environment###

#data$names <- factor(data$names , levels=c("A", "D", "C", "B"))
#use the above line to re-order the data in how you would like it presented in the bar chart format

total.boxplot.facet <- norm.data %>%
    select(status , s.v) %>%
    pivot_longer(., cols = c(s.v), names_to = "Var", values_to = "Val") %>%
    ggplot(aes(x = Var, y = Val, fill = status)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=0.7, notch = F, notchwidth = 0.1, varwidth = FALSE) +
    #The notch displays a confidence interval around the median which is normally based on the median +/- 1.58*IQR/sqrt(n).                 Notches are used to compare groups; if the notches of two boxes do not overlap, this is a strong evidence that the medians differ.
    #geom_jitter(size = 0.1, shape=16, position=position_jitter(0.2)) +
    #changes color of the fill
    ggtitle("INSERT_TITLE_HERE") +
    ylab("") + 
    xlab("Adjusted p value (FDR) cutoff (p<0.05)") +
    theme(text = element_text(face = "bold", size = 4.0)) +
    scale_fill_manual(values=c("#808080", "#E69F00", "#56B4E9","#660000","#6666FF","#0000FF","#000099","#000066"))+
    #Manually changes the color of the variables 
    #https://www.rapidtables.com/web/color/RGB_Color.html
    #see above website for good tool to find color
    theme_bw() +
    #manually change the theme of the plot (use google for more ideas)
    stat_summary(fun = mean, geom="point",colour="red1", size=1, position=position_dodge(width=0.75)) +
    #change the position_dodge width if the mean points are not aligning correctly with the groups
    theme(legend.title = element_blank()) +
    #changes or deletes  legend title size--change too element_text(size = XX)
    theme(legend.text =  element_text(size = 10)) +
    #changes legend text size 
    theme(legend.position='bottom') +
    # deletes entire legend
    theme(axis.text=element_text(size=10), axis.title=element_text(size=14,face="bold")) +
    #changes the size of axis text (first), and then size of axis labels (second)
    theme(legend.background = element_rect(fill= NULL, size=0.5, linetype="solid", colour ="black")) +
    #creates box around legend 
    theme(axis.ticks.x = element_blank()) +
    facet_wrap(~ Var, nrow = 4,  labeller = labeller(groupwrap = label_wrap_gen(1000)), scales = "free") +
    #facet wraps --> set number of rows here also specifies the width of the facets
    theme(axis.text.x=element_blank()) +
    theme(plot.title = element_text(size=22)) +
    #change sizew of title
     theme(strip.text.x = element_text(size = 10, color = "Black", face = "bold")) + 
    #changes the size of the facet plot titles and changes the font style face = bold.italic is also good
    ylim(-2.0, 2.0) + 
    #changes y limits for size 
    geom_signif()
#need to fix the wrap here and also set up parameters for the width/ height


#my_comparisons <- list(c("Dart","Dropnet"), c("Dropnet","Helo"), c("Dart", "Helo"))
#old code that is somehow still here -- not really quite sure what waas supossed to be happening here

#total.boxplot.facet <- total.boxplot.facet + stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
#stat_compare_means(label.y = 50)     # Add global p-value


total.boxplot.facet




ggsave("High_Quality_boxplot.pdf", plot = total.boxplot.facet, units="in", width = 20, height = 16, dpi=5000)

#note a 5 row ANOOVA is width= 26 and height 14

#change to .tiff if extra HQ file is needed

#ggsave("High_Quality_boxplot.tiff", plot = total.boxplot.facet, units="in", width = 18, height = 12, dpi=1000, compression = 'lzw')
#change to .tiff and use compression +'lzw' in ggsave


get_legend(total.boxplot.facet, position = "right")

leg <- get_legend(total.boxplot.facet)
leg <- as_ggplot(leg)

ggsave("Legend_here.pdf", plot = leg, units = "in", width= 10, height= 10, dpi= 2000 )
#prints just the legend for figure creation if needed

```


```{r}
dodge <- position_dodge(width = 1.0)
#aligns the boxplots (below) with the violin and can adjust the spacing between the plots 

total.violin <- norm.data %>%
    select(status , s.v) %>%
    pivot_longer(., cols = c(s.v), names_to = "Var", values_to = "Val") %>%
    ggplot(aes(x = Var, y = Val, fill = status)) +
    geom_violin(trim= FALSE, position = dodge) +
    #position is set above on line 179...for boxplots to align later 
    #geom_jitter(size = 1.4) + 
    #turn on the above to be able to see the jitter overlaid onto the plot
    scale_fill_manual(values=c("#808080", "#E69F00", "#56B4E9","#660000","#6666FF","#0000FF","#000099","#000066"))+
    ggtitle("INSERT_TITLE_HERE") +
    ylab("") + 
    xlab("Adjusted p value (FDR) cutoff (p<0.05)") +
    theme_bw() +
    theme(text = element_text(face = "bold", size = 8.5)) +
    scale_fill_manual(values=c("#808080", "#E69F00", "#56B4E9","#660000","#6666FF","#0000FF","#000099","#000066"))+
    #Manually changes the color of the variables 
    #https://www.rapidtables.com/web/color/RGB_Color.html
    #see above website for good tool to find color
    stat_summary(fun = mean, geom="point",colour="red1", size=1, position=position_dodge(width=0.75)) +
    #change the position_dodge width if the mean points are not aligning correctly with the groups
    theme(legend.title = element_blank()) +
    #changes or deletes  legend title size--change too element_text(size = XX)
    theme(legend.text =  element_text(size = 25)) +
    #changes legend text size 
    theme(legend.position='bottom') +
    # deletes entire legend
    theme(axis.text=element_text(size=20), axis.title=element_text(size=25,face="bold")) +
    #changes the size of axis text (first), and then size of axis labels (second)
    theme(legend.background = element_rect(fill= NULL, size=0.5, linetype="solid", colour ="black")) +
    #creates box around legend 
    theme(axis.ticks.x = element_blank()) +
    facet_wrap(~ Var, nrow = 3,  labeller = labeller(groupwrap = label_wrap_gen(15)), scales = "free") +
    #facet wraps --> set number of rows here also specifies the width of the facets
    theme(axis.text.x=element_blank()) +
    
    theme(plot.title = element_text(size=30)) +
    #change sizew of title
     theme(strip.text.x = element_text(
        size = 20, color = "Black", face = "bold"
        )) + 
    #changes the size of the facet plot titles
    ylim(-3.5, 3.5)
    #changes y limits for size 

total.violin <- total.violin + geom_boxplot(width=0.1, position = dodge)
#adds boxplots inside the violin plots 


#total.violin

ggsave("High_Quality_Violin.pdf", plot = total.violin, units="in", width= 28, height= 24, dpi=5000)
#ggsave("High_Quality_Violin.tiff", plot = total.violin, units="in", width= 6, height=5, dpi=400, compression = 'lzw')

#exports jpeg


###CODE AT BOTTOM WILL MAKE SINGLE METABOLITE PLOT ...NEEDS TO BE ACTIVATED!

```

```{r}

library(gt)
library(glue)
library(data.table)

all_mean_FC <- aggregate(norm.data[], list(norm.data$status), mean)
all_mean_FC <- t(all_mean_FC)
#this code will create table with the means for all the FC calculated earlier

all_mean_FC <- all_mean_FC[-c(1,2), ]  
#deletes two rows 

options(digits = 3)

all_mean_FC <- as.data.frame(all_mean_FC)
#saves as data frame


setDT(all_mean_FC, keep.rownames = TRUE)[]
#makes row names into column and turns to data.table
names(all_mean_FC)[names(all_mean_FC) == "rn"] <- "Metabolite"
names(all_mean_FC)[names(all_mean_FC) == "V1"] <- "Dart"
names(all_mean_FC)[names(all_mean_FC) == "V2"] <- "Dropnet"
names(all_mean_FC)[names(all_mean_FC) == "V3"] <- "Helicopter"
#changes names of the columns 

all_mean_FC <- all_mean_FC[-c(1,2), ]  
#deletes two rows 

#is.num <- sapply(all_mean_FC, is.numeric)
#all_mean_FC[is.num] <- lapply(all_mean_FC[is.num], round, 3)
#tries to round the results--not working as of 8/28/20...also only want to do this because of the table 



##MAKE THIS A GT TABLE 
gt_tbl <- gt(data = all_mean_FC)

# Show the gt Table
gt_tbl <-gt_tbl %>%
  tab_header(
    title = "Title_Here",
    subtitle = "Mean Fold Changes Between Metabolites"
  )

gt_tbl <- 
  gt_tbl %>%
  tab_source_note(
    source_note = "One Way ANOVA, Adjusted p-value (FDR) cutoff:0.05, Tukey's post-hoc analysis "
  )


# Show the gt Table
gt_tbl
#still need to write code to export the 
gtsave(gt_tbl, filename = "Full_ANOVA_FC_Mean_Table.pdf")

```

#this was done to calculate the specific differences between the groups for ONE specific metabolite and to visualize it

status <- as.data.frame(norm.data$status)

single.metabolite <- as.data.frame(norm.data$Urea)
#change the specific metabolite here

combined.data <- cbind(status, single.metabolite)

names(combined.data)[names(combined.data) == "norm.data$status"] <- "Capture"
names(combined.data)[names(combined.data) == "norm.data$Urea"] <- "Metabolite"

example_metab <-aggregate(combined.data[, 2], list(combined.data$Capture), mean)
#above code will show specific metabolite
example_metab

is.num <- sapply(example_metab, is.numeric)
example_metab[is.num] <- lapply(example_metab[is.num], round, 3)


#creates ggplot2 objects
FC_barplot <- ggplot(data=example_metab, aes(x=Group.1, y=x, fill = Group.1)) +
                geom_bar (stat = "identity") +
                geom_text(aes(label=x), vjust=-0.3, size=2.0)+
                theme_minimal() +
                ggtitle("Mean (Insert_metabolite) Fold Change") +
                ylab("Fold Change (FC)") + 
                xlab("Adjusted p value (FDR) cutoff (p<0.05)") +
                theme(text = element_text(face = "bold", size = 4.0)) +
                scale_fill_manual(values=c("#1B9E77", "#7570B3", "#D95F02"))+
                #Manually changes the color of the variables 
                #https://www.rapidtables.com/web/color/RGB_Color.html
                #see above website for good tool to find color
                theme(legend.title = element_blank()) +
                #deletes legend title
                theme(legend.position='none') +
                # deletes entire legend
                theme(axis.text=element_text(size=10), axis.title=element_text(size=14,face="bold")) +
                #changes the size of axis text (first), and then size of axis labels (second)
                #theme(legend.background = element_rect(fill= NULL, size=0.5, linetype="solid", colour ="black")) +
                #creates box around legend 
                #facet wraps --> set number of rows here also specifies the width of the facets
                theme(plot.title = element_text(size=22)) 
                #change sizew of title
                #ylim(-2, 2)
                #changes y limits for size 
            #FC_barplot
            
ggsave("Single_Metabolite_ANOVA.tiff", plot = FC_barplot, units="in", width= 6, height=5, dpi=1000, compression = 'lzw')
